### 2025-07-30 paper/figures/corr_table.tex
cut lots

\begin{tabular}[t]{l>{}r>{}r>{}r>{}r>{}r}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{5}{c}{Correlation with differential harm and:} \\
\cmidrule(l{3pt}r{3pt}){2-6}
State & Exp. Dem. seats & Efficiency gap & Partisan bias & Mean-median & Declination\\
\midrule
Utah & \cellcolor[HTML]{8A671F}{$-0.983$} & \cellcolor[HTML]{75689E}{$0.957$} & \cellcolor[HTML]{927421}{$-0.893$} & \cellcolor[HTML]{8E6C20}{$-0.945$} & \cellcolor[HTML]{947722}{$-0.876$}\\
Mississippi & \cellcolor[HTML]{9D8627}{$-0.772$} & \cellcolor[HTML]{BFD9D5}{$0.207$} & \cellcolor[HTML]{E8EEE3}{$0.031$} & \cellcolor[HTML]{E7E5C0}{$-0.096$} & \cellcolor[HTML]{A89735}{$-0.656$}\\
Iowa & \cellcolor[HTML]{AD9F3D}{$-0.597$} & \cellcolor[HTML]{A0B5C2}{$0.449$} & \cellcolor[HTML]{DED99D}{$-0.185$} & \cellcolor[HTML]{C4DCD7}{$0.184$} & \cellcolor[HTML]{A9C2C9}{$0.365$}\\
North Carolina & \cellcolor[HTML]{8E6C20}{$-0.943$} & \cellcolor[HTML]{8A88AE}{$0.740$} & \cellcolor[HTML]{9AAABC}{$0.520$} & \cellcolor[HTML]{8C8DB0}{$0.704$} & \cellcolor[HTML]{9BAABC}{$0.519$}\\
Wisconsin & \cellcolor[HTML]{998024}{$-0.816$} & \cellcolor[HTML]{929BB4}{$0.621$} & \cellcolor[HTML]{C2DBD7}{$0.191$} & \cellcolor[HTML]{9096B2}{$0.652$} & \cellcolor[HTML]{A8C1C8}{$0.374$}\\
New Jersey & \cellcolor[HTML]{9B8425}{$-0.789$} & \cellcolor[HTML]{9FB4C1}{$0.459$} & \cellcolor[HTML]{DCE8DF}{$0.082$} & \cellcolor[HTML]{EEF0E0}{$-0.014$} & \cellcolor[HTML]{A7C0C7}{$0.381$}\\
Massachusetts & \cellcolor[HTML]{987E23}{$-0.826$} & \cellcolor[HTML]{E9E9CA}{$-0.069$} & \cellcolor[HTML]{E1DDA7}{$-0.158$} & \cellcolor[HTML]{CADFD9}{$0.159$} & \cellcolor[HTML]{808080}{NA}\\
California & \cellcolor[HTML]{AEA03E}{$-0.590$} & \cellcolor[HTML]{939BB4}{$0.616$} & \cellcolor[HTML]{DCD798}{$-0.199$} & \cellcolor[HTML]{D1C97B}{$-0.302$} & \cellcolor[HTML]{AAC3C9}{$0.359$}\\
\bottomrule
\end{tabular}